---
title: "Linear Regression"
author: "Oliver Imhans"
date: "2022-12-21"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE, fig.height = 3.3, 
fig.width = 5.5, fig.align = "center")
```

For convenience, I will add all the necessary libraries at the very beginning.
```{r}
library(tidyverse)
library(dplyr)
library(corrgram)
library(DataExplorer)
library(ppcor)
library(caTools)
library(ggplot2)
library(corrplot) 
library(data.table)
library(plotly)
library(lm.beta)
```


In mathematics, regression is a statistical technique that is employed when the relationship between dependent variables and independent variables is considered. This process is used to determine if the changes in the dependent variables are connected with any of the independent variables.

# Linear regression

This is the most commonly used type of predictive analysis. In simple terms, this is a linear (arranged along a straight line) approach for relationship modeling between two variables. The variables are always **dependent** and **independent**. It is important to note that the order of the variables matters. The independent variable belongs on the x-axis, while the dependent variable belongs on the y-axis.
There are two types of linear regression:\ 

i. Simple Linear Regression\
ii. Multiple Linear Regression

The linear regression for two variables is based on the linear equation $y = mx + c$ where m and c are constants. 

The graph of a linear equation of the form above is a *straight line*. 

**Example** 
Plot the graph of $y = 2x + 6$ using the range -3 to 3. 

```{r}
# First define the equation
lin_equ <- function(x){2*x+6}

# Then plot the equation
ggplot(data.frame(x=c(-3, 3)), aes(x=x)) + 
  stat_function(fun=lin_equ)
```
**How to perform Simple Linear Regression**

The formula for linear regression is $y= b_1 X + b_0$. But in a more standard form, the complete linear regression model is:\
$$ y = b_1X + b_0 + \epsilon$$
where:

y is the predicted value 

$b_0$ is the intercept.

$b_1$ is the regression coefficient 

X is the independent variable 

$\epsilon$ is the error of the estimate.

The aim of linear regression is to find the line of best fit that goes through the data set. This is achieved by searching for $b_1$ the regression coefficient that will minimize the $\epsilon$ the error of the model.   



In the world of Data Science, *linear regression is an algorithm* that predicts the outcome from the linear relationship between the independent variables and dependent variables. From the foregoing, linear regression is classified as a supervised learning algorithm. 
There are some benefits to using linear regression\
1. It is easily scalable.\
2. It is easily implemented.\
3. It is relatively straightforward.

The example below will focus on how to use R to create a regression model. **R linear regression uses the lm() function to create this regression model.**
To view this model, the summary() function will be used. 


The dataset for this example is available at the link: https://www.kaggle.com/datasets/karthickveerakumar/salary-data-simple-linear-regression?resource=download

```{r}
# Importing the dataset
data = read.csv('salary.csv')
```

Taking a quick overview of the data
```{r}
summary(data)
```


**A quick visualization of the data**

```{r}
# scatter plot 
ggplot(data, aes(x=YearsExperience, y=Salary)) + 
    geom_point()
```



```{r}
# Splitting the data into the training data and testing data

split = sample.split(data$Salary, SplitRatio = 0.8)
train_data = subset(data, split == TRUE)
test_data = subset(data, split == FALSE)
```



```{r}
# Fit the Linear Regression model into the training data
model = lm(formula = Salary ~ YearsExperience,
                        data = train_data)
coef(model)
```
**NB** The results show the intercept and the beta coefficient for the YearsExperience variable.

# Interpretation

From the output above:

i. The estimated regression line equation can be written as follow: $$Salary = 27266.90 + (9243.10 * YearsExperience)$$

ii. The intercept $(b_0)$ is 27266.90. It can be interpreted as the predicted salary  for a zero YearsExperience. 

The linear model for this dataset is built and a formula that can be used to make a prediction is also derived. Before using this regression model, I have to ensure that it is statistically significant. 


```{r}
# model summary
summary(model)  
```

```{r}
# Predicting the Test set results
ypredict = predict(model, newdata = test_data)
```


```{r}
actual_predict <- data.frame(cbind(actual=test_data$Salary, predicted=ypredict))  
head(actual_predict)
```


**Checking for statistical significance**\
i.    This model can be considered statistically significant only when both the p-Values are less that the pre-determined statistical significance level, which is ideally 0.05. This is visually interpreted by the significant stars at the end of the row. The more stars beside the variable’s p-Value, the more significant the variable.
```{r}
reg_p <- function (pvalue) 
  {if (class(pvalue) != "lm") stop("Not an object of class 'lm' ")
   f <- summary(pvalue)$fstatistic
   p <- pf(f[1],f[2],f[3],lower.tail=F)
   attributes(p) <- NULL
   return(p)}

reg_p(model)
```
This value corresponds with model summary

ii.   R-Squared: the higher the better $(> 0.70)$\
```{r}
summary(model)$r.squared 
```
iii.    Adj R-Squared\: the higher the better\
```{r}
summary(model)$adj.r.squared
```

iv.   F-Statistic: the higher the better\
```{r}
summary(model)$fstatistic 
```
v. Std. Error: When it is closer to zero the better.\
```{r}
modelSummary <- summary(model)  
modelCoeffs <- modelSummary$coefficients  
beta.estimate <- modelCoeffs["YearsExperience", "Estimate"]  
std.error <- modelCoeffs["YearsExperience", "Std. Error"] 
std.error
```

vi. t-statistic: this	should be greater than 1.96 for p-value to be less than 0.05. \
```{r}
t_value <- beta.estimate/std.error  
t_value
```

vii.    AIC	Lower the better \
```{r}
AIC(model)
```

viii.   BIC	Lower the better \
```{r}
BIC(model)
```

ix.   MAPE (Mean absolute percentage error): The lower the better \
```{r}
mape <- mean(abs((actual_predict$predicted - actual_predict$actual))/actual_predict$actual)
mape
```

x.   MSE (Mean squared error)	Lower the better \
```{r}
mean(model$residuals^2)
```


xi.    Min_Max Accuracy => mean(min(actual, predicted)/max(actual, predicted))	The higher the better
```{r}
min_max_accuracy <- mean(apply(actual_predict, 1, min) / apply(actual_predict, 1, max))  
min_max_accuracy
```




**Visualization**
```{r}
# Visualizing the training data outcome 
ggplot() + geom_point(aes(x = train_data$YearsExperience,
                y = train_data$Salary), colour = 'red') +
geom_line(aes(x = train_data$YearsExperience,
y = predict(model, newdata = train_data)), colour = 'blue') +
         
ggtitle('Salary vs Experience (Train Data)') +
xlab('Years of experience') +
ylab('Salary')
```

```{r}
# Visualizing the test data outcome 
  ggplot() +
  geom_point(aes(x = test_data$YearsExperience, y = test_data$Salary),
             colour = 'red') +
  geom_line(aes(x = train_data$YearsExperience,
             y = predict(model, newdata = train_data)),
             colour = 'blue') +
  ggtitle('Salary vs Experience (Test Data)') +
  xlab('Years of experience') +
  ylab('Salary')
```


# Finding Accuracy

```{r}
rmse<-sqrt(mean(ypredict -data$YearsExperience)^2)
rmse
```

**Multiple Linear Regression**

Multiple linear regression is an extension of simple linear regression used to predict an outcome variable (y) on the basis of multiple distinct predictor variables (x).

With three predictor variables (x), the prediction of y is expressed by the following equation:

$$y = b_0 + b_1*X_1 + b_2*X_2 + b_3*X_3$$
**Below is an example**
```{r}
multi_data<-data("marketing", package = "datarium")
head(marketing, 4)
```

Visualization
```{r}
ggplot(marketing, aes(x = facebook+youtube+newspaper, y = sales)) +
  geom_point() +
  stat_smooth()
```
```{r}
cor(marketing$sales, marketing$facebook)
```


```{r}
# Building the model
multi_model <- lm(sales ~ youtube + facebook + newspaper, data = marketing)
summary(multi_model)
```

```{r}
# Regression line
ggplot(marketing, aes(youtube + facebook + newspaper, sales)) +
  geom_point() +
  stat_smooth(method = lm)
```




Residual Standard Error (RSE), or sigma
```{r}
sigma(multi_model)/mean(marketing$sales)
```

**Conclusion**





